---
title: "Assignment 9: Sentiment Analysis of Emma"
author: "Anna Rescifina"
format: html
execute:
  echo: true
  warning: false
  message: false
---

## Overview 

In this assignment, I use the **tidytext** framework and three sentiment lexicons (**nrc**, **afinn**, and **bing**) to explore positive and joyful words in *Emma* by Jane Austen. I first prepare and tokenize the text, then join it with each lexicon, count frequent sentiment words, and visualize the results with bar charts and word clouds.

## SetUp and Libraries Loading

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo       = TRUE, 
                      fig.align  = "center",
                      fig.height = 3, fig.width = 4)
ggplot2::theme_set(ggplot2::theme_bw() + ggplot2::theme(strip.background = ggplot2::element_rect(fill = "white")))

library(tidyverse)
library(tidytext)
library(janeaustenr)
library(stringr)
library(wordcloud)
library(RColorBrewer)
```

## Step 1: Data wrangling 

In this step I load the 'Jane Austen' dataset and prepare the data for analysis. I create a "linenumber" variable to track each line and a "chapter" variable by detecting chapter headings using a regular expression. Finally, I filter the dataset to keep only the book 'Emma'.

```{r}
emma_lines <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(
      text, 
      regex("^chapter [\\divxlc]", ignore_case = TRUE)
    ))
  ) %>%
  ungroup() %>%
  filter(book == "Emma")
```

## Step 2: Tokenization 

In this step I convert the text into a tidy format using "unnest_tokens()", where each word becomes a separate row. This format is required for sentiment analysis.

```{r}
emma_words <- emma_lines %>%
  unnest_tokens(word, text)
```

## Step 3: NRC Sentiment Analysis (Joy)

Using the NRC lexicon I focus specifically on words associated with the emotion joy. I join the tokenized words with the NRC dataset, count word frequencies, and keep only the most frequent words (n\>50).

```{r}
emma_nrc <- emma_words %>%
  inner_join(
    get_sentiments("nrc") %>% filter(sentiment == "joy"),
    by = "word"
  ) %>%
  count(word, sort = TRUE) %>%
  filter(n > 50)
```

## Visualization (Bar Chart)

```{r}
emma_nrc %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Joy Words in Emma (NRC)", 
    x = "Frequency", 
    y = NULL 
  )
```

## Word Cloud 

```{r}
wordcloud(
  words = emma_nrc$word,
  freq = emma_nrc$n,
  min.freq = 1,
  max.words = 100,
  random.order = FALSE,
  colors = brewer.pal(8, "Dark2")
)
```

## Step 3b: AFINN Sentiment Analysis (Positive)

The AFINN lexicon assigns numeric sentiment scores. In this step, I select only words with positive sentiment (Score\>= 1); then count and visualize the most frequent ones.

```{r}
emma_afinn <- emma_words %>%
  inner_join(
    get_sentiments("afinn") %>% filter(value >= 1),
    by = "word"
  ) %>%
  count(word, sort = TRUE) %>%
  filter(n > 50)
```

## Visualization (Bar Chart)

```{r}
emma_afinn %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(fill = "darkgreen") +
  labs( 
    title = "Positive Words in Emma (AFINN)",
    x = "Frequency", 
    y = NULL
    )
```

## Word Cloud 

```{r}
wordcloud(
  words = emma_afinn$word,
  freq = emma_afinn$n,
  min.freq = 1,
  max.words = 100,
  random.order = FALSE,
  colors = brewer.pal(8, "Set2")
)
```

## Step 3c: BING Sentiment Analysis (Positive)

The BING lexicon classifies words as either positive or negative. Here, I filter for positive words, count their frequency, and visualize the most common ones.

```{r}
emma_bing <- emma_words %>%
  inner_join(
    get_sentiments("bing") %>% filter(sentiment == "positive"),
    by = "word"
  ) %>%
  count(word, sort = TRUE) %>%
  filter(n > 50)
```

## Visualization (Bar Chart)

```{r}
emma_bing %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(x = n, y = word)) +
  geom_col(fill = "purple") +
  labs(
    title = "Positive Words in Emma (BING)", 
    x = "Frequency",
    y = NULL
  )
```

## Word Cloud 

```{r}
wordcloud(
  words = emma_bing$word,
  freq = emma_bing$n,
  min.freq = 1,
  max.words = 100,
  random.order = FALSE,
  colors = brewer.pal(8, "Paired")
)
```

## Conclusion 

In this analysis I used three different sentiment lexicons to identify positive and joyful words in *Emma*. While each lexicon defines sentiment differently, all three highlight recurring positive language in the novel. The visualizations show that *Emma* contains a consistent presence of emotionally positive vocabulary.
